---
title: "mbta eda"
author: "xiaoyanbin Cai"
date: "2022-12-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## loading libraries
```{r}
library(shiny)
library(googleway)
library(dplyr)
library(leaflet)
library(knitr)
library(tidyverse)
library(lubridate)
```

## import data
```{r}
traveltime_1 <- read.csv("~/Desktop/615-final-/2022-Q1_HRTravelTimes.csv")
head(traveltime_1)

traveltime_1.2 <- read.csv("~/Desktop/615-final-/2022-Q1_LRTravelTimes.csv")
head(traveltime_1.2)
```

## Extract one week data 
```{r}
traveltime_1$service_date <- as.Date(traveltime_1$service_date)
data1 <- traveltime_1[traveltime_1$service_date>='2022-01-01' & traveltime_1$service_date<='2022-01-07',]
data1
```

## Extract Jan data
```{r}
traveltime_1.2$service_date <- as.Date(traveltime_1.2$service_date)
data1.2 <- traveltime_1.2[traveltime_1.2$service_date>='2022-01-01' & traveltime_1.2$service_date<='2022-01-07',]
data1.2
```

## Put two data set together
```{r}
total_1 = dplyr::bind_rows(data1,data1.2)
total_1

#summary(total)
total_1 <- total_1 %>% mutate(route_id = factor(route_id))
glimpse(total_1)
```

## import data
```{r}
traveltime_2 <- read.csv("~/Desktop/615-final-/2022-Q1_HRTravelTimes.csv")
head(traveltime_2)

traveltime_3 <- read.csv("~/Desktop/615-final-/2022-Q1_LRTravelTimes.csv")
tail(traveltime_3)
```

## extract Fab data
```{r}
traveltime_2$service_date <- as.Date(traveltime_2$service_date)
data2 <- traveltime_2[traveltime_2$service_date>='2022-02-01' & traveltime_2$service_date<='2022-02-07',]
tail(data2)
```

## extarct March data
```{r}
traveltime_3$service_date <- as.Date(traveltime_3$service_date)
data3 <- traveltime_3[traveltime_3$service_date>='2022-03-01' & traveltime_3$service_date<='2022-03-07',]
data3
tail(data3)
```

## extarct April data
```{r}
traveltime_4 <- read.csv("~/Desktop/615-final-/2022-Q2_HRTravelTimes.csv")
head(traveltime_4)
traveltime_4.2 <- read.csv("~/Desktop/615-final-/2022-Q2_LRTravelTimes.csv")
head(traveltime_4.2)
```

## put 2 data together
```{r}
traveltime_4$service_date <- as.Date(traveltime_4$service_date)
data4 <- traveltime_4[traveltime_4$service_date>='2022-04-01' & traveltime_4$service_date<='2022-04-07',]
tail(data4)

traveltime_4.2$service_date <- as.Date(traveltime_4.2$service_date)
data4.2 <- traveltime_4.2[traveltime_4.2$service_date>='2022-04-01' & traveltime_4.2$service_date<='2022-04-07',]
tail(data4.2)
```

```{r}
total_4 = dplyr::bind_rows(data4,data4.2)
total_4
```

## extract May data
```{r}
traveltime_4$service_date <- as.Date(traveltime_4$service_date)
data5 <- traveltime_4[traveltime_4$service_date>='2022-05-01' & traveltime_4$service_date<='2022-05-07',]
tail(data5)

traveltime_4.2$service_date <- as.Date(traveltime_4.2$service_date)
data5.2 <- traveltime_4.2[traveltime_4.2$service_date>='2022-05-01' & traveltime_4.2$service_date<='2022-05-07',]
tail(data5.2)
```

```{r}
total_5 = dplyr::bind_rows(data5,data5.2)
total_5
```

## extract June data

```{r}
traveltime_4$service_date <- as.Date(traveltime_4$service_date)
data6 <- traveltime_4[traveltime_4$service_date>='2022-06-01' & traveltime_4$service_date<='2022-06-07',]
tail(data6)

traveltime_4.2$service_date <- as.Date(traveltime_4.2$service_date)
data6.2 <- traveltime_4.2[traveltime_4.2$service_date>='2022-06-01' & traveltime_4.2$service_date<='2022-06-07',]
tail(data6.2)
```

```{r}
total_6 = dplyr::bind_rows(data6,data6.2)
total_6
```

## extract July data

```{r}
traveltime_7 <- read.csv("~/Desktop/615-final-/2022-Q3_HRTravelTimes.csv")
head(traveltime_7)
traveltime_7.2 <- read.csv("~/Desktop/615-final-/2022-Q3_LRTravelTimes.csv")
head(traveltime_7.2)
```

```{r}
traveltime_7$service_date <- as.Date(traveltime_7$service_date)
data7 <- traveltime_7[traveltime_7$service_date>='2022-07-01' & traveltime_7$service_date<='2022-07-07',]
tail(data7)

traveltime_7.2$service_date <- as.Date(traveltime_7.2$service_date)
data7.2 <- traveltime_7.2[traveltime_7.2$service_date>='2022-07-01' & traveltime_7.2$service_date<='2022-07-07',]
tail(data7.2)
```


## put two data set together
```{r}
total_7 = dplyr::bind_rows(data7,data7.2)
total_7
```

## Put 2022 data set all together
```{r}
total_2022 = dplyr::bind_rows(total_1,data2,data3,total_4,total_5,total_6,total_7)
total_2022
```

## Trip durations and their distributions
```{r}
sample <- sample_n(total_2022,100000)#randomly choose 100000 data from total data set
plot1 <- ggplot(sample)+geom_histogram(binwidth =100, aes(travel_time_sec))+ ggtitle("Fig 1.trip durations and their distribution") + ylab ("count") + xlab("trip duration")
plot1
```

## Trip durations and their distributions transformation
```{r}
sample <- sample_n(total_2022,10000000)#randomly choose 100000 data from total data set
plot2 <- ggplot(sample)+geom_histogram(aes(log(travel_time_sec)))+ ggtitle("Fig 2.trip durations and their distribution after transformation") + ylab ("frequenct of trips") + xlab("log trip duration")+ xlim(2,10)
plot2 
```

## To evaluate how travel time vary over months
```{r}
total_2022 %>%
  group_by(service_date) %>%
  summarise(mean_travel = mean(travel_time_sec)) %>%
  arrange(desc(mean_travel))
```
* As you can see from the result, on 2022-02-02 there's highest average travel time of subway from 2022 MBTA data frame.


```{r}
service_date_1 <- format(as.Date(total_2022$service_date), "%Y/%m")
service_date_1
ggplot(total_2022,aes(x=factor(service_date_1),y=travel_time_sec)) + geom_boxplot() + labs(x="date",y="travel time")
```
* This boxplot has given us a clear vision about the most reliable measure and mean.

```{r}
ggplot(data=total_2022,aes(x=route_id,fill = service_date_1)) + geom_bar()
```



